\begin{acknowledgements}      

    \newcommand{\tnk}{\textsc}

    \looseness -1 \lettrine{T}{his} thesis would never have been possible without the
    ideas, friendship, support, and assistance of numerous people. Foremost among them is,
    of course, \tnk{Xavier Martorell}, who is an infinitely patient advisor, caring and
    understanding. His immense faith in my abilities and our research group is shown in
    the diverse research topics we could explore. And yet, on each of these topics, he was
    a constant source of ideas and inspiration, and I feel enormously privileged to have
    him as my advisor. In addition to his undying enthusiasm and energy, he also directed
    me to numerous people to generate a wide range of scientific contacts that made this
    journey successful.
    
    \looseness -1 I am grateful to \tnk{Marc Gonzalez} for the initial insights on
    building the power and performance models in the early stages of my PhD.  His guidance
    and willingness to engage in discussions made the initial proceedings of this thesis
    possible. 
    
    \looseness -1 I  owe a lot of thanks to \tnk{Paul M. Carpenter}. Firstly, for guiding
    me to produce high quality results and for lending me a book on writing
    style~\citep{paulbook}. I cannot thank him enough for his generosity and willingness
    to help in the hard times of my PhD i.e., with multiple paper rejections and not knowing
    what to do.  The amount of time spent together discussing problems (be it late
    evenings or mid afternoon), brainstorming ideas, and comprehending some concepts from
    text books was invaluable in understanding the causal-effect relationships presented
    in chapter~\ref{chapter: hipster}. This PhD would have been harder without Paul
    teaching me -- consciously and unconsciously -- how to understand a problem in an
    abstract way.  It was a great pleasure having him as my advisor at Barcelona
    Supercomputing Center.
    
    For over half a decade, I have had the pleasure to work with \tnk{Daniel Moss\'e} who
    taught me how to sketch out an idea and design experiments. His dedication and
    resilience to do excellent research, by focusing on each data point and each aspect of
    this, is shown in section~\ref{subsection: evaluation repp-h}.
    
    \looseness -1 I especially owe much to \tnk{Vinicius Petrucci},  \tnk{Tomasz Patejko}
    and \tnk{Pieter Bellens}, who took time during and after numerous cups of coffee, for
    the countless days of brainstorming (with their humorous comments), correcting papers
    and showing me how to rewrite a scientific article.  Being a slow learner, they taught
    me twice.  Many atimes my crazy ideas bounced off them and came back to me looking
    \emph{even more crazy}. We had a lot of fruitful discussions and I enjoyed working
    with them.
    
    \looseness -1 My friends in Barcelona (now everywhere) made my life both in and
    outside of work engaging and enjoyable. These people are (in alphabetical order)
    \tnk{Alfonso Rodr\'iguez}, \tnk{Amandine Iseux}, \tnk{Joanna Hoddinott}, \tnk{Kallia
    Chronaki}, \tnk{Mary Ocampo}, \tnk{Rahul Gayatri}, \tnk{Rommel Sanchez}, \tnk{Sicong
    Zhuang} and \tnk{Siddharth Bhaskara}.  Although we did not get a chance to work
    together, their support and patience to listen to all my failures and joy of
    paper acceptances was considerate.  They always cheered me up after a disastrous
    evening with, a long bicycle ride, good food and deep philosophical conversation.

    \looseness -1 Since I am half-way into my Academic Awards speech here anyway, I also
    want to thank my parents \tnk{Padmavathi} and \tnk{Anand Kumar} \tnk{Nishtala}.  Ever
    since the beginning of elementary school you have always insisted in me the
    importance of education (even if that meant pulling me from underneath the bed!) and
    raised me with the love and importance of science and math from an early age.
    \tnk{Amma}, I will always remember you waking me up every single day 
    during my stay away from home and for teaching me to be patient even in hard times.
    \tnk{Nanna}, I clearly remember you taking time from your busy schedule to learn
    and teach me to program in BASIC and LOGO at the age of 7 -- your dedication and work
    ethic are truly inspiring. Along with my sister \tnk{Dr. Nirupama Nishtala}, my family
    continued to believe in me even when I did not. I hope I made them proud.
    
    \looseness -1 Last, but certainly not least, I want to thank my best friend,
    \tnk{Aleksandra Dokurno} for her moral support, for listening to all my problems, and
    to each day's success and failures.  Her astonishing drive to pay meticulous attention
    to detail is contagious, and without her, I might still be the one without ``any''
    academic papers! I would also like to thank her for staying up with me -- at times --
    on the long nights of paper writing and giving me amazing food. Thank you for what you
    have done to me.
    
    Cheers to y'all, you made this ride memorable!  
    
    \vspace{2mm}
    
    \textit{This work has been partially supported by the European Union FP7 program
    through the Mont-Blanc-2 (FP7-ICT-610402) project, by the Ministerio de Economia y
    Competitividad under contract Computaci\'on de Altas Prestaciones VII
    (TIN2015-65316-P), and the Departament de Innovaci\'o, Universitats i Empresa de la
    Generalitat de Catalunya, under project MPEXPAR: Models de Programaci\'o i Entorns d
    Execuci\'o Paral$\cdot$lels (2014-SGR-1051).}
    

\end{acknowledgements}
