\chapter*{Publication List}

\noindent \looseness -1 This thesis is based in part on the following published papers:

\begin{enumerate}[leftmargin=*] 
    
    \looseness -1 \item[{\small {\circled{1}}}] \textbf{Rajiv Nishtala}, Marc Gonzalez
        Tallada, Xavier Martorell, ``\textsf{A Methodology to Build Models and Predict
        Performance-Power in CMPs}'', in the 44th International Conference on Parallel
        Processing Workshops ({ICPPW}), Beijing, China, Sep. 1-4, 2015.
    
    \looseness -1 \item[{\small {\circled{2}}}] \textbf{Rajiv Nishtala}, Xavier Martorell,
        Vinicius Petrucci, Daniel Moss\'e, ``\textsf{REPP-H: Runtime Estimation of Power
        and Performance on Heterogeneous Data Centers}'', in the 28th International
        Symposium on Computer Architecture and High Performance Computing ({SBAC-PAD}),
        Los Angeles, USA, Oct. 26-28, 2016.

    \looseness -1 \item[{\small {\circled{3}}}] \textbf{Rajiv Nishtala}, Xavier Martorell,
        ``\textsf{RePP-C: Runtime Estimation of Performance-Power with Workload
        Consolidation in CMPs}'', in the 7th International Green and Sustainable Computing
        Conference ({IGSC}), Hangzhou, China, Nov. 7-9, 2016. 
    
    \looseness -1 \item[{\small {\circled{4}}}] \textbf{Rajiv Nishtala}, Paul Carpenter,
        Vinicius Petrucci, Xavier Martorell, ``\textsf{Hipster: Hybrid Task Manager for
        Latency-Critical Cloud Workloads}'', in the 23rd IEEE Symposium on High
        Performance Computer Architecture ({HPCA}) 2017, Austin, USA, Feb. 4-8, 2017.
    

\end{enumerate} 

\noindent \looseness -1 The following publication is under review:

\begin{enumerate}[leftmargin=*] 
    

    \item[\small {\circled{1}}] \textbf{Rajiv Nishtala}, Paul Carpenter, Vinicius
        Petrucci, Xavier Martorell, ``\textsf{The Hipster Approach for Improving Cloud System Efficiency}''
        in ACM Transactions of Computer Systems ({ACM} {TOCS}) Feb. 2017.

\end{enumerate}

\noindent \looseness -1 The following paper was my masters' work, and not included in this thesis:

\begin{enumerate}[leftmargin=*]  

    \looseness -1 \item[{\small {\circled{1}}}] \textbf{Rajiv Nishtala}, Daniel Moss\'e,
        Vinicius Petrucci, ``\textsf{Energy-aware thread co-location in heterogeneous
        multicore processors}'', in the 11th ACM International Conference on Embedded
        Software ({EMSOFT}), Montreal, Canada, Sep. 29 - Oct. 4, 2013.
    
\end{enumerate}
